### 2020-05-16 slides/slides.tex

\documentclass[english,aspectratio=43,t]{beamer}
% useful options:
%   aspectratio=169

\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{listings}
\usepackage[iso]{isodate}

% select theme before using any of its features
\usetheme{Comsys}

\title[Cool]{The Origins of Congestion\\and Network assisted end-to-end\\Congestion Control}
\author{Til Mohr}
\authorextra{Advisor: Jan Rüth}
\subtitle{Proseminar}
\date{\today}
%\location{Aachen}
%\projectlogo[scale=0.5]{logos/rwth_comsys_cropped}
%\logoextra{logos/rwth_comsys_cropped}

%\titlegraphics{images/presentation}
%\begin{titlepicture}
%	\includegraphics[width=3cm]{images/presentation}
%\end{titlepicture}

\begin{document}

\titleframe

\begin{frame}
	\frametitle{Outline}
	\tableofcontents
\end{frame}

\section{First Topic}
\label{sec:first_topic}
\frame\sectionpage

\begin{frame}{Blocks}
	\begin{cols}{0.5,auto}
		\begin{block}{Block with Title}
			Content of block with title
		\end{block}
		\begin{whiteblock}{White Block}
			Just a blue frame around the text
		\end{whiteblock}
		\begin{exampleblock}{This is an Example Block}
			Example Block
		\end{exampleblock}
		\begin{alertblock}{This is an Alert Block}
			Alert Block
		\end{alertblock}
	\col
		\begin{block}{}
			Content of block without title
		\end{block}
		\begin{whiteblock}{}
			Without title
		\end{whiteblock}
		\begin{grayblock}{}
			Grayed out block
		\end{grayblock}
		\begin{exampleblock}{}
			Example Block without Title
		\end{exampleblock}
		\begin{redblock}{}
			Red Block without Title
		\end{redblock}
	\end{cols}
	\pushdown
	\begin{alertblock}{}
		\begin{center}Don't overload your slides, beware of the shadow!\end{center}
	\end{alertblock}
\end{frame}

\begin{frame}{Itemize}
	\begin{cols}{1,2*1}
		\begin{itemize}
		\item item
		\item item
		\item []
		\item item
			\begin{itemize}
			\item subitem
				\begin{itemize}
				\item subsubitem
				\end{itemize}
			\end{itemize}
		\end{itemize}
	\col
		\begin{enumerate}
		\item Enumerated item
		\item Enumerated item
		\item[]
		\item Enumerated item
		\item Enumerated item
		\end{enumerate}
	\end{cols}

	\vspace{1em}

	\begin{cols}{auto}
		\begin{block}{Block with Items}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{block}
		\begin{block}{}
			\begin{enumerate}
			\item item
			\item item
			\end{enumerate}
		\end{block}
	\col
		\begin{exampleblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{exampleblock}
		\begin{alertblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{alertblock}
	\col
		\begin{whiteblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{whiteblock}
		\begin{grayblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{grayblock}
	\end{cols}
\end{frame}

\begin{frame}{Illustrative Pictures right of Main Text}
	\begin{itemize}
	\item A picture can be inserted which is right aligned
		\begin{itemize}
		\item It is vertically aligned with the position\ilgraphics[height=2em]{logos/rwth}
		\\it was inserted in the source
		\end{itemize}
	\item There are two variants
		\begin{itemize}
		\item A variant based on \textit{{\textbackslash}includegraphics} which takes the same arguments as \textit{{\textbackslash}includegraphics}
		\item An environment for any content
		\begin{illustration}
			\begin{tikzpicture}
				\draw[draw=red] (0,0) circle (1em);
				\draw[draw=blue] (0,0) rectangle(2em,2em);
			\end{tikzpicture}
		\end{illustration}
		\end{itemize}
	\end{itemize}
\end{frame}

\singlemessageframe{This Frame contains\\only a single Message!}

\begin{frame}{Sizes}%
	\showsizes[cm]%
	\showfonts
\end{frame}

\end{document}